### 2017-10-17 paper/sections/1_introduction.tex
paper: draft 0
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% What is the problem?
File system metadata services in HPC have scalability problems. Administrative
tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the file
system~\cite{zheng:pdsw2014-batchfs} lead to contention for the same
directories and inodes ({\it e.g.}, path traversal). Applications perform
better with dedicated metadata servers~\cite{sevilla:sc15-mantle,
ren:sc2014-indexfs} but provisioning a metadata server for every client is
unreasonable. This problem is exacerbated by current trends in HPC, where
architectures are transitioning from complex storage stacks with burst buffer,
file system, object store, and tape tiers to more simplified stacks with just a
burst buffer and object store~\cite{bent:login16-hpc-trends}; this puts more
pressure on data access because more requests end up hitting the same layer and
latencies cannot be hidden while data migrates across tiers.

\begin{figure}[tb]
\centering
\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}
\caption{ Clients that need better performance can relax consistency on their
subtree ({e.g.}, HDFS) or even decouple the subtree and move it locally
({e.g.}, BatchFS, RAMDisk). Decoupled subtrees can further relax durability
semantics to achieve even better performance. Meanwhile, clients that require
stronger guarantees ({\it e.g.,} POSIX) can still reside in the same namespace.
}\label{fig:subtree-policies}
\end{figure}

@@

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improves
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). We also show the effect of having a metadata specific
file format in systems that are based on in-memory data structures. Because
this paper focuses on file system metadata protocols, we only scale up to 20
clients executing an extremely metdata intensive workload; this load is the
maximum capacity of a single metadata server for the platform we chose to
implement Cudele on.  Partitioning and replicating metadata across a cluster
can be controlled with something like Mantle~\cite{sevilla:sc15-mantle}, but in
this paper we show the capacity and performance of a metadata server with
superior metadata protocols, which should be used to inform load balancing
decisions.

>>>>

% What is the problem?
File system metadata services in HPC have scalability problems because
administrative tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the
file system~\cite{zheng:pdsw2014-batchfs}, contend for the same directories and
inodes. Applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
metadata server for every client is unreasonable. This problem is exacerbated
by current trends in HPC, where architectures are transitioning from complex
storage stacks with burst buffer, file system, object store, and tape tiers to
more simplified stacks with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
because more requests end up hitting the same layer and latencies cannot be
hidden while data migrates across tiers.

\begin{figure}[tb]
\centering
\includegraphics[width=0.35\textwidth]{figures/subtree-policies1.png}

\caption{For performance, clients can relax consistency on their subtree
({e.g.}, HDFS) or even decouple the subtree and move it locally ({e.g.},
BatchFS, RAMDisk). Decoupled subtrees can further relax durability semantics
for even better performance. Clients that require stronger guarantees ({\it
e.g.,} POSIX) can still reside in the same namespace.
}\label{fig:subtree-policies}

\end{figure}

@@

% Results
Our results confirm the assertions of ``clean-state" research systems that
decouple namespaces; specifically that the technique drastically improves
performance (104\(\times\) speed up) but we go a step further by quantifying
the costs of merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down). We also show the effect of having a metadata specific
file format in systems that are based on in-memory data structures.  We scale
up to 20 clients executing an extremely metadata intensive job because this
workload exhausts the resources of a single metadata server for the codebase we
built on.  Partitioning and replicating metadata across a cluster can be
controlled with something like Mantle~\cite{sevilla:sc15-mantle}, but in this
paper we show the capacity and performance of a metadata server with superior
metadata protocols, which should be used to inform load balancing decisions.

### 2017-10-17 paper/sections/3_posix_overheads.tex
paper: draft 0
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
      \caption{Runtime normalized to no journal} \label{fig:overhead-a}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
      \caption{Interference forces \texttt{lookup()}s}
      \label{fig:overhead-b}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere.png}
      \caption{Interference hurts variability}
      \label{fig:overhead-c}
  \end{subfigure}
  \caption{The overhead of global durability and strong consistency in CephFS.
  (a) shows the effect of different journal segment sizes, which are streamed
  into the object store for fault tolerance. (b) and (c) show that when a second
  client ``interferes", capabilities are revoked and metadata servers do more
  work.  \label{fig:overhead}}
\end{figure*}

@@

% benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:overhead-b}.  If only one client is creating files in a
directory (``isolated" curve on \(y1\) axis) then that client can lookup the
existence of new files locally before issuing a create request to the metadata
server. If another client starts creating files in the same directory
(``interfere" curve on \(y1\) axis) then the directory inode transitions out of
read caching and the first client must send \texttt{lookup()}s to the metadata
server (``interfere" curve on \(y2\) axis). These extra requests increase the
throughput of the ``interfere" curve because the metadata server can handle the
extra load but performance suffers.  This degradation is shown in
Figure~\ref{fig:overhead-c}, where we scale the number of clients and show
increased variability. The results are normalized to a single
isolated client and the \(y\) axis is the standard deviation of the runtime of
all clients.  For the ``interfere" bars, each client creates files in private
directories and at 30 seconds we launch another process that creates files in
those directories.  For less than 7 clients, the runtime and deviation is worse
when clients interfere. At 7 clients, the metadata server is overloaded so the
directory caching in the ``isolated" bars has no benefit.\\

>>>>

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
      \caption{Runtime normalized to no journal} \label{fig:overhead-a}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
      \caption{Interference forces \texttt{lookup()}s}
      \label{fig:overhead-b}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere-scale.png}
      \caption{Interference hurts variability}
      \label{fig:overhead-c}
  \end{subfigure}
  \caption{The overhead of global durability and strong consistency in CephFS.
  (a) shows the effect of different journal segment sizes, which are streamed
  into the object store for fault tolerance. (b) and (c) show that when a second
  client ``interferes", capabilities are revoked and metadata servers do more
  work.  \label{fig:overhead}}
\end{figure*}

@@

% benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:overhead-b}.  If only one client is creating files in a
directory (``isolated" curve on \(y1\) axis) then that client can lookup the
existence of new files locally before issuing a create request to the metadata
server. If another client starts creating files in the same directory
(``interfere" curve on \(y1\) axis) then the directory inode transitions out of
read caching and the first client must send \texttt{lookup()}s to the metadata
server (``interfere" curve on \(y2\) axis). These extra requests increase the
throughput of the ``interfere" curve because the metadata server can handle the
extra load but performance suffers.  This degradation is shown in
Figure~\ref{fig:overhead-c}, where we scale the number of clients and show the
slowdown of the slowest client.  The results are normalized to a single
isolated client without a metadata server journal.  For the ``interfere" bars,
each client creates files in private directories and at 30 seconds we launch
another process that creates files in those directories. 18 clients is the
maximum number the metadata server could handle for this metadata-intensive
workload; at higher client load, the metadata server started complaining about
laggy and unresponsive requests.\\

### 2017-10-17 paper/sections/6_results.tex
paper: draft 0
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\noindent\textbf{Takeaway}: the Cudele mechanisms have overheads and costs
that can differ {\it by orders of magnitude}. Cudele gives users the ability
to compose these mechanisms based on their application's correctness
requirements and performance goals.

@@

Clients creating files in private directories is heavily
studied in HPC~\cite{plfs, cephfs04, indexfs, giga+, batchfs, deltafs,
sevilla:eurosys, shardfs, mantle}, mostly due to checkpoint-restart~\cite{}.
But the workload also appears in cloud workloads, where systems like Hadoop use
the file abstraction to exchange work units to workers or to indicate when jobs
complete~\cite{}. A more familiar example is uncompressing an archive ({\it
e.g.}, \texttt{tar xzf}), where the file system services a flash crowd of
creates across all directories as shown in Figure~\ref{fig:overhead-creates}.
We use this as a microbenchmark because it allows us to control the size of the
log, since each create creates a single journal event.

@@

% Final file system metadata states are equivalent No guarantees while
% transitioning mechanisms
We compare these semantics because the final metadata states are equivalent.
Cudele makes no guarantees during execution of the mechanisms or when
transitioning semantics -- the semantics are guaranteed {\it once the mechanism
completes}. So if servers fail during a mechanism, metadata or data may be
lost.\\

% DeltaFS vs. BatchFS vs. POSIX: decoupled is faster
% (5-7x) vs. (90-104x) slower
% Does not scale?
% Strong Consistency > 10X slower than Weak Consistency 
{\it Speedups of Decoupled Namespaces:} Weak consistency uses the
decoupled namespace strategy and shows up to a \(20\times\) speedup over the
traditional namespaces that use RPCs. Compared to the baseline the slowdown is
\(5-7\times\) for Strong Consistency, which emulates BatchFS and
\(90-104\times\) for Weak Consistency, which emulates DeltaFS.\\

% POSIX (no stream) vs. POSIX: cost of durability < consistency
{\it Durability \(<<\) Consistency:} The \(1.15\times\) overhead of
Strong Consistency compared to Strong Consistency, No Fault Tolerance for
100K files is negligible. It suggests that the overhead of consistency is much
larger than the overhead of durability. This conclusion should be stronger as
we scale the number of files because the cost of streaming the journal into the
object store is constant. We omit the same analysis for Weak Consistency
because the runtimes are so short that the normalized slowdowns are misleading.\\

@@

\subsubsection{Network Bandwidth}
% 1G - 10G
\subsubsection{Number of Clients}
% up to 30 clients
\subsubsection{Number of Serialized Merges}

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
      \caption{Merging journals serially.} \label{fig:merge-a}
  \end{subfigure}\\
  ~
  \begin{subfigure}[b]{\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/merge-b.png}
      \caption{Merging a journal w/ 100K create updates.}
      \label{fig:merge-b}
  \end{subfigure}
  \caption{Merging updates has linear scalability.\label{fig:merge}}
\end{figure}

Cudele serializes merge requests at the metadata server before applying the
updates to the in-memory metadata store. Merging concurrently is an
optimization for future work, but merges for the largest journal size we tested
were quite fast.  Furthermore, since we place no restrictions on the validity
of metadata inserted into the journal, we can avoid touching poorly scaling
data structures. In other words, we can scale linearly if we weaken our
consistency by never checking for the existence of files before creating files.

Figure~\ref{fig:merge-a} demonstrates this scalability by showing the runtime
(\(y\) axis) of merging journals of different sizes (\(x\) axis). The variance
of the runtime is over 20 serial merges, each hitting a different directory in
the namespace.  The runtime scales linearly and with low variance because the
``Create" mechanism that was used to create the journal only appends
\texttt{open()} requests to the journal. When the updates are merged by the
metadata server into the in-memory metadata store they never scan growing data
structures. Had we implemented the ``Create" mechanism to include
\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
the poor scaling that we see with the ``RPCs" mechanism.
Figure~\ref{fig:merge-b} confirms that we do not touch a poorly scaling data
structure because the runtimes do not increase as we add more metadata. \\

@@

\subsection{Use Case 2: Creates with Rogue Client}
Clients create files in private directories and a separate client, which we
call a ``rogue" client, creates files in each directory. This introduces false
sharing and the metadata server revokes capabilities on directories touched by
the rogue client. While HPC tries to avoid these situations with
worklflows~\cite{batchfs, deltafs}, it stills happens in distributed file
systems when users unintentionally access directories in a shared file system.

@@

>>>>

\noindent\textbf{Takeaway}: the Cudele mechanisms have overheads and costs
that can differ {\it by orders of magnitude}. Cudele gives users the ability
to compose these mechanisms based on their application's correctness
requirements and performance goals.\\

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
\caption{ [\href{https://...}{source}] Merging journals serially.}
\label{fig:merge}
\end{figure}

%\begin{figure}[t]
%  \centering
%  \begin{subfigure}[b]{\linewidth}
%      \centering
%      \includegraphics[width=1.0\linewidth]{graphs/merge-a.png}
%      \caption{Merging journals serially.} \label{fig:merge-a}
%  \end{subfigure}\\
%  ~
%  \begin{subfigure}[b]{\linewidth}
%      \centering
%      \includegraphics[width=1.0\linewidth]{graphs/merge-b.png}
%      \caption{Merging a journal w/ 100K create updates.}
%      \label{fig:merge-b}
%  \end{subfigure}
%  \caption{Merging updates has linear scalability.\label{fig:merge}}
%\end{figure}

{\it Number of Serialized Merges:} Cudele serializes merge requests at the
metadata server before applying the updates to the in-memory metadata store.
Merging concurrently is an optimization for future work, but merges for the
largest journal size we tested were quite fast.  Furthermore, since we place no
restrictions on the validity of metadata inserted into the journal, we can
avoid touching poorly scaling data structures. In other words, we can scale
linearly if we weaken our consistency by never checking for the existence of
files before creating files.

Figure~\ref{fig:merge} demonstrates this scalability by showing the runtime
(\(y\) axis) of merging journals of different sizes (\(x\) axis). The variance
of the runtime is over 20 serial merges, each hitting a different directory in
the namespace.  The runtime scales linearly and with low variance because the
``Create" mechanism that was used to create the journal only appends
\texttt{open()} requests to the journal. When the updates are merged by the
metadata server into the in-memory metadata store they never scan growing data
structures. Had we implemented the ``Create" mechanism to include
\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
the poor scaling that we see with the ``RPCs" mechanism.\\
%Figure~\ref{fig:merge-b} confirms that we do not touch a poorly scaling data
%structure because the runtimes do not increase as we add more metadata. \\

@@

% CITEME xiao:socc2015-shardfs CITEME HADOOP
Clients creating files in private directories is heavily studied in
HPC~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs, patil:fast2011-giga,
zheng:pdsw2014-batchfs, sevilla:sc15-mantle}, mostly due to
checkpoint-restart~\cite{bent_plfs_2009}.  But the workload also appears in
cloud workloads, where systems like Hadoop use the file abstraction to exchange
work units to workers or to indicate when jobs
complete~\cite{shvachko:login2012-hdfs-scalability}. A more familiar example is
uncompressing an archive ({\it e.g.}, \texttt{tar xzf}), where the file system
services a flash crowd of creates across all directories as shown in
Figure~\ref{fig:overhead-creates}.  We use this as a microbenchmark because it
allows us to control the size of the log, since each create creates a single
journal event.

@@

% Final file system metadata states are equivalent No guarantees while
% transitioning mechanisms
We compare these semantics because the final metadata states are equivalent.
Cudele makes no guarantees during execution of the mechanisms or when
transitioning semantics -- the semantics are guaranteed {\it once the mechanism
completes}. So if servers fail during a mechanism, metadata or data may be
lost.

% DeltaFS vs. BatchFS vs. POSIX: decoupled is faster
% (5-7x) vs. (90-104x) slower
% Does not scale?
% Strong Consistency > 10X slower than Weak Consistency 
{\it Speedups of Decoupled Namespaces:} Weak consistency uses the
decoupled namespace strategy and shows up to a \(20\times\) speedup over the
traditional namespaces that use RPCs. Compared to the baseline the slowdown is
\(5-7\times\) for Strong Consistency, which emulates BatchFS and
\(90-104\times\) for Weak Consistency, which emulates DeltaFS.

% POSIX (no stream) vs. POSIX: cost of durability < consistency
{\it Durability \(<<\) Consistency:} The \(1.15\times\) overhead of
Strong Consistency compared to Strong Consistency, No Fault Tolerance for
100K files is negligible. It suggests that the overhead of consistency is much
larger than the overhead of durability. This conclusion should be stronger as
we scale the number of files because the cost of streaming the journal into the
object store is constant. We omit the same analysis for Weak Consistency
because the runtimes are so short that the normalized slowdowns are misleading.

@@

@@

\subsection{Use Case 2: Creates with Rogue Client}

Clients create files in private directories and a separate client, which we
call a ``rogue" client, creates files in each directory. This introduces false
sharing and the metadata server revokes capabilities on directories touched by
the rogue client. While HPC tries to avoid these situations with
worklflows~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}, it stills
happens in distributed file systems when users unintentionally access
directories in a shared file system.

@@

% CITEME
Scientists use the file system to check the progress of jobs using
\texttt{ls}~\cite{CITEME}. The number of files or size of the files is
indicative of the progress. This practice is not too different from systems
that use the file system to manage the progress of jobs; Hadoop writes to
temporary files, renames them when complete, and creates a ``DONE" file to
indicate to the runtime that the task did not fail and should not be
re-scheduled on another node. In this scenario, Cudele users will not see the
progress of decoupled namespace since their updates are not globally visible.
To help scientists judge the progress of their jobs, Cudele has a ``namespace
sync" that sends batches of updates back to the global namespace at regular
intervals.

\subsection{Use Case 4: Reading Large Directories}

% CITEME
The final uses case is reading large directories. At job completion, scientists
might use \texttt{ls} again to see the results. This causes great strain on the
file system as paths need to traversed and the entire directory, with all its
entries, must be transferred back to the client. Here we show how decoupling a
large namespace and materializing the view in memory on the client is faster
than doing RPCs for walks of the file system namespace.
